Cartogram Mapping and its Application to Cancer Data Visualisation

### 2019-09-10 review_article.Rmd
Australian cancer atlas animation

```{r choropleth-grid-create, message=FALSE, fig.cap = "Public facing choropleth cancer maps.[a. The Environment and Health Atlas of England and Wales](http://www.envhealthatlas.co.uk/eha/Breast/) (Lung cancer in females)  (Incidence rates per 100,000 for all cancers, male and female),  [b. Bowel Cancer Australia Atlas](http://www.bowelcanceratlas.org/), [c. Atlas of Cancer in QLD - Relative incidence ration, Lung cancer in males](https://cancerqld.org.au/research/queensland-cancer-statistics/queensland-cancer-atlas/), [d. Globocan 2012: Estimated  Cancer Incidence, Mortality and Prevalence](http://globocan.iarc.fr/Pages/Map.aspx), [e. United States Cancer Statistics: An Interactive Cancer Statistics Website](https://nccd.cdc.gov/DCPC_INCA/) (Incidence rate per 100,000, all cancer types, men and women), [f. Map of Cancer Mortality Rates in Spain](http://elpais.com/elpais/2014/10/06/media/1412612722_141933.html) (Relative risk, Lung cancer, men vs women), [g. An Atlas of Cancer in South Australia](https://www.cancersa.org.au/assets/images/pdfs/An%20Atlas%20of%20Cancer%20in%20South%20Australia%20-%20Full%20Report.pdf) (Incidence rate per 100,000, lung cancer in females) "}
library(cowplot)

# Import the images
plot1 <- png::readPNG('cancer_map_images/choropleths/21.choropleth.png')
plot2 <- png::readPNG('cancer_map_images/choropleths/15.choropleth.png')
plot3 <- png::readPNG('cancer_map_images/choropleths/25.choropleth.png')
plot4 <- png::readPNG('cancer_map_images/choropleths/3.choropleth.png') 
plot5 <- png::readPNG('cancer_map_images/choropleths/7.choropleth.png')
plot6 <- png::readPNG('cancer_map_images/choropleths/23.choropleth.png')
plot7 <- png::readPNG('cancer_map_images/choropleths/26.choropleth.png')

ggdraw() +
  draw_plot(rasterGrob(plot1), 0, .75, 0.33, .25) +
  draw_plot(rasterGrob(plot2), 0.33, 0.75, 0.33, .25) +
  draw_plot(rasterGrob(plot3), 0.66, 0.5, .33, .5) +
  draw_plot(rasterGrob(plot4), 0, 0.5, .33, .25) +
  draw_plot(rasterGrob(plot5), 0.33, 0.5, .33, .25) +
  draw_plot(rasterGrob(plot6), 0, 0, .66, .5) +
  draw_plot(rasterGrob(plot7), .66, 0, .33, .5) +
  draw_plot_label(c("a", "b", "c", "d", "e", "f", "g"), 
    c(0.015, 0.33, 0.63, 0.015, 0.33, 0.015, 0.63), 
    c(0.99, 0.99, 0.99, 0.74, 0.74, 0.49, 0.49), size = 15)

@@

|  | Choropleth | Contiguous | Non-contiguous | Dorling | Tilegrams |
|:-----------:|------------|-----------|---------------|---------|-----------|
| Feature preservation: |  |  |  |  |  |
| Preserves shapes | Y | N | Y | N | N |
| Preserves neighbours | Y | Y | Y | S | S |
| Uniform use of shape | N | N | N | Y | Y |
| Bivariate display* | N | Y | Y | Y | Y |
* A bivariate display refers to a map that allows two variables to be plotted. One variable via the size of the area, and another by filling the area with a colour.

@@

<!-- Animations -->
@CPISACA provide weather maps as well developed and appropriate examples of spatial displays to communicate to the general public, and encourage those creating disease maps to leverage the success of weather displays.
The movement of a weather system will follow a forecasted path, the animation will communicate this so that all map users can follow the path of the weather system across the geography over a specified period of time.

The move from static, print media displays toward interactivity allow users to develop maps to ask and answer their own questions.
Developmets and adoptions of this technology has allowed map users more control of the message displayed.
Users are able to see the familiar shapes, zoom in to explore densely populated communities, or change the display completely to highlight the population space.
@CPISACA provide three questions that are generally when using a disease map:

@@

<!-- Interactivity-->
This oppportunity provides room to move away from the traditional approach of a single map for a single use, into a more explorative territory.
Interactivity allows users to drive their own exploration, controlling a display through variable controls, links, tool tips, and data selection tools [@TGA].

>>>>

```{r choropleth-grid-create, message=FALSE, fig.cap = "Public facing choropleth cancer maps.[a. The Environment and Health Atlas of England and Wales](http://www.envhealthatlas.co.uk/eha/Breast/) (Lung cancer in females)  (Incidence rates per 100,000 for all cancers, male and female), [b. Globocan 2012: Estimated  Cancer Incidence, Mortality and Prevalence](http://globocan.iarc.fr/Pages/Map.aspx),  [c. Atlas of Cancer in QLD - Relative incidence ration, Lung cancer in males](https://cancerqld.org.au/research/queensland-cancer-statistics/queensland-cancer-atlas/), [d. Bowel Cancer Australia Atlas](http://www.bowelcanceratlas.org/), [e. United States Cancer Statistics: An Interactive Cancer Statistics Website](https://nccd.cdc.gov/DCPC_INCA/) (Incidence rate per 100,000, all cancer types, men and women), [f. Map of Cancer Mortality Rates in Spain](http://elpais.com/elpais/2014/10/06/media/1412612722_141933.html) (Relative risk, Lung cancer, men vs women), [g. An Atlas of Cancer in South Australia](https://www.cancersa.org.au/assets/images/pdfs/An%20Atlas%20of%20Cancer%20in%20South%20Australia%20-%20Full%20Report.pdf) (Incidence rate per 100,000, lung cancer in females) "}
library(cowplot)

# Import the images
plot1 <- png::readPNG('cancer_map_images/choropleths/21.choropleth.png')
plot2 <- png::readPNG('cancer_map_images/choropleths/3.choropleth.png') 
plot3 <- png::readPNG('cancer_map_images/choropleths/25.choropleth.png')
plot4 <- png::readPNG('cancer_map_images/choropleths/15.choropleth.png')
plot5 <- png::readPNG('cancer_map_images/choropleths/7.choropleth.png')
plot6 <- png::readPNG('cancer_map_images/choropleths/23.choropleth.png')
plot7 <- png::readPNG('cancer_map_images/choropleths/26.choropleth.png')

ggdraw() +
  draw_plot(rasterGrob(plot1), 0, .75, 0.33, .25) +
  draw_plot(rasterGrob(plot2), 0.33, 0.75, 0.33, .25) +
  draw_plot(rasterGrob(plot3), 0.66, 0.5, .33, .5) +
  draw_plot(rasterGrob(plot4), 0, 0.5, .33, .25) +
  draw_plot(rasterGrob(plot5), 0.33, 0.5, .33, .25) +
  draw_plot(rasterGrob(plot6), 0, 0, .66, .5) +
  draw_plot(rasterGrob(plot7), .66, 0, .33, .5) +
  draw_plot_label(c("a", "b", "c", "d", "e", "f", "g"), 
    c(0.015, 0.33, 0.63, 0.015, 0.33, 0.014, 0.63), 
    c(0.99, 0.99, 0.99, 0.74, 0.74, 0.49, 0.49), size = 15)

@@

|  | Choropleth | Contiguous | Non-contiguous | Dorling | Tilegrams |
|:-----------:|------------|-----------|---------------|---------|-----------|
| Preserves shapes | Y | N | Y | N | N |
| Preserves neighbours | Y | Y | Y | S | S |
| Uniform use of shape | N | N | N | Y | Y |
| Bivariate display^[A bivariate display refers to a map that allows two variables to be plotted. One variable via the size of the area, and another by filling the area with a colour.] | N | Y | Y | Y | Y |

@@

<!-- Animations -->
@CPISACA provide weather maps as well developed and appropriate examples of spatial displays to communicate to the general public, and encourage those creating disease maps to leverage the success of weather displays.
The movement of a weather system will follow a forecasted path, the animation will communicate this so that all map users can follow the path of the weather system across the geography over a specified period of time.
In the disease mapping space, the Australian Cancer Atlas [-@TACA] provides [tours](
https://atlas.cancer.org.au/app/tour/lungcancer) which change the display to draw user's attention to areas on the map that are relevant to the story. This directs user's attention and gives them tools to plan their own exploration.

@CPISACA provide three questions that are generally when using a disease map:

@@

<!-- Interactivity-->
The move from static, print media displays toward interactivity allow users to develop maps to ask and answer their own questions.
Developmets and adoptions of this technology has allowed map users more control of the message displayed.
Users are able to see the familiar shapes, zoom in to explore densely populated communities, or change the display completely to highlight the population space.

This oppportunity provides room to move away from the traditional approach of a single map for a single use, into a more explorative territory.
Interactivity allows users to drive their own exploration, controlling a display through variable controls, links, tool tips, and data selection tools [@TGA].
